---
title: "Climate Data Analysis"
author: "Imran"
date: "`r Sys.Date()`"
output: html_document
---

## 📌 Introduction
This report analyzes global temperature changes over time using the **GlobalTemperatures.csv** dataset.  
The report includes:  
- A **line chart** showing global temperature trends.  
- A **scatter plot** of CO₂ emissions vs. temperature anomaly.  

---

```{r setup, include=FALSE}
# Load necessary libraries
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
library(ggplot2)
library(dplyr)

# Set working directory (modify accordingly)
setwd("C:\\Users\\Student\\Desktop\\PFAI_LUFFY\\OEL\\T2") 

# Load the dataset
data <- read.csv("GlobalTemperatures.csv")

# Convert dt column to Date format
data$dt <- as.Date(data$dt, format="%Y-%m-%d")

# Extract Year
data$Year <- as.numeric(format(data$dt, "%Y"))

# Remove rows with NA values inLandAverageTemperature
data <- data %>% filter(!is.na(LandAverageTemperature))

# Aggregate yearly average temperature
yearly_data <- data %>%
  group_by(Year) %>%
  summarise(AverageTemperature = mean(LandAverageTemperature, na.rm = TRUE))

# Display first few rows
head(data)

# Check for missing values
colSums(is.na(data))

# Ensure data is not empty
if (nrow(yearly_data) == 0) {
  stop("Error: No valid temperature data available after filtering.")
}

# Create line chart
ggplot(yearly_data, aes(x = Year, y = AverageTemperature)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Global Temperature Change Over Time", 
       x = "Year", y = "Temperature (°C)") +
  theme_minimal()


# Assuming a column for CO2 emissions exists, replace 'CO2_Emissions' with actual column name
if ("CO2_Emissions" %in% colnames(data)) {
  ggplot(data, aes(x = CO2_Emissions, y = LandAverageTemperature)) +
    geom_point(color = "red", alpha = 0.5) +
    labs(title = "CO₂ Emissions vs. Temperature Anomaly", 
         x = "CO₂ Emissions", y = "Temperature (°C)") +
    theme_minimal()
} else {
  print("No CO₂ emissions data available in the dataset.")
}

```{r plot_temperature, echo=FALSE, fig.width=8, fig.height=5}
# Ensure data is not empty
if (nrow(yearly_data) > 0) {
  print(
    ggplot(yearly_data, aes(x = Year, y = AverageTemperature)) +
      geom_line(color = "blue", size = 1) +
      labs(title = "Global Temperature Change Over Time", 
           x = "Year", y = "Temperature (°C)") +
      theme_minimal()
  )
} else {
  print("No valid data available for plotting.")
}
